\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-04-05 redaction/dhp_july2017.tex
ajustement
test
changement d'un mot pour voir si j'arrive à puller
changement texte

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of export price multiplied by the ad-valorem cost (expressed in Norwegian crowns), i.e. $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results). One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them, allowing us to gain in generality in this respect. Making use of our estimates for 2004 (see tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.5\% in air transport. This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases). In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment. \citet{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs. For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$. Given our estimates for 2004 (see Table \ref{tab:result_air_3d_detail} in Appendix \ref{app:more_results}) such that $x = 0.028$ and $0.015$ in vessel and air transport respectively, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport. On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs. In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs trend patterns.

\textbf{a ajuster et remonter figure annexe.}

Two main comments can be made. First, in both transport modes, the additive component appears of sizeable importance all over the period. Confirming the results obtained in Section \ref{sec:results_decomposition}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon. By contrast, they represent an sizeable and structural dimension of international transport costs. However, one can also notice that, if the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport over the period, it also substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered (see Figure \ref{fig:part_cout_additif}, Appendix \ref{app:more_results}).\smallskip

@@

In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database. This focus is shared with part of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}. Specifically, we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects, in the same spirit as \cite{hummels2007}.

@@

As a first step, Figure \ref{fig:decomp_TC_3d} displays the \textbf{total transport costs (the total of the two, expressed in percentage of the fas price)}, by transport mode for each year between 1974 and 2013.

\begin{figure}[htbp]
\caption{Trend in Transport costs (Yearly mean value, 3 digits, in \% of the fas price, 1989 exclude for "Air")}
\label{fig:decomp_TC_3d}
\begin{center}
\includegraphics[width=6.5in, height=5in]{Figure1_Trend_of_total_TC_by_mode.pdf}
\end{center}
\end{figure}

\textbf{figure a changer et texte a adapter}

Figure \ref{fig:decomp_TC_3d} reports that both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport. \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates. In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:decomp_TC_3d} springs from high oil-shock related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture. The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs. We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} On US data, \citet{Hummels_1999} obtains that overall transport cost declined from 6 \% to 4 \% of the import value between 1974 and 1996. For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping. Overall, our results display trends that are close to those reported by \citet{Hummels_1999}.\smallskip

Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade. Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased , independently of any change in transport costs \textit{per se}. As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs. Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974. Our own results tend to challenge this view, as we further show.

@@

\paragraph{Challenging the $\beta$'s constancy assumption} Our results of Section \ref{sec:results_decomposition} (on a year-to-year basis, see Appendix \ref{app:more_results}) and of Figure \ref{fig:part_cout_additif}, cast some doubt on the empirical relevance of this assumption. As reported in Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension), the share of additive costs is rather varying over time, in both transport modes. In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}.

We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs. To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, i.e. $\beta$, according to:
\begin{eqnarray*}
\beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}-1},
\end{eqnarray*}

@@

Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$). Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect. Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974. We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:decomp_TC_3d}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).

@@

\paragraph{Obtaining the unfitted total transport cost index} For each transport mode, we build the ``total'' transport cost based on Equation (\ref{eq:base_estimee}) according to:
$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}. Recall that $\widehat{\tau}^{adv}_t-1$ measuring the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed in percentage of the fas price. We then transform this value in an index with basis year 1974, applying a similar formula as above (by transport mode):
$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$

@@

\subsection{Characterizing the time trends in transport costs}
Figure \ref{fig:totalTC_compeffects_excl} reports the results.\footnote{Figure \ref{fig:totalTC_compeffects_excl} reports the results, all types of goods included. In Appendix \ref{app:comp-effects}, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.} Panels (a), (b) and (c), we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for Air transport (starting from the reference value 100 in 1974). Panels (d), (e) and (f) report the results for Vessel. In each panel, we report the evolution of transport costs for both the unfitted (plain blue line) and the fitted (dotted red line) measures.

@@

Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}. First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)). International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping. This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}). Second, and as discussed in Section \ref{sec:results_decomposition}, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)). Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s. Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs. Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl}, we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the ``ceteris paribus'' transport costs in Air transport. For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period (Figure \ref{fig:totalTC_compeffects_excl}, panels (a) to (c)). Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the ``ceteris paribus'' transport costs. Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), where they amplify the reduction in ``pure'' transport costs (the plain line being lower than the dotted line). Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs \textit{per se} (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two). This is particularly the case for the multiplicative component (panel (d)).\smallskip

@@

These results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the ``per se'' or ``ceteris paribus'' transport costs for both air and maritime shipping. If anything, we find the opposite result here. This drives us to investigate further this difference. As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension. Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap). As direct consequence, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions. %By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the ``ceteris paribus'' transport cost dimension, as we show below. To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004). The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}. The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of ``pure'' transport costs, i.e. composition effects excluded ($\widehat{TC}_t$).

@@

These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}. Applying \citeh{hummels2007} method, we obtain that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as pointed in panels (c) and (f) from Figure \ref{fig:totalTC_compeffects_excl}, panels c and f.

Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs. In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in vessel transport. Complementing the findings of Section \ref{sec:results_decomposition}, these results point out the importance of integrating the additive dimension of international transport costs, here in view of characterizing their time trends.

\subsection{Time trends in transport costs: Robustness to the weighting scheme \label{sec:robustness}}

As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the ``per se'' or ``ceteris paribus'' transport costs over time. In this section, we provide a robustness test to this difference in the weighting scheme implemented to extract the trade composition effects.\smallskip

When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow. Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974. In this section, our objective is to assess that this difference of treatment is not responsible for the difference of results relative to the role of the trade composition effects, that we rather attribute to the modeling of the additive component.

@@

Our results could be extended in two main ways. On the empirical side, one may want to go deeper in the ``structural'' determinants of trade costs, i.e. identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices. On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) and in affecting the international transmission of business cycles. This is left for further research.

@@

\paragraph{More on additive costs} In Figure , we report the share of additive costs in total transport costs over time in both Air and Vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}, conditional on a year and a transport mode). Precisely, each point in time is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent (i.e., in percent of the export price $\widetilde{p}$).

\begin{figure}[htbp]
\caption{Share of additive costs}
\label{fig:part_cout_additif}
\begin{center}
\includegraphics[width=10cm, height=7cm]{part_cout_additif.pdf}
%{\footnotesize {OECD data} }
\end{center}
\end{figure}

@@

In this section, we detail our methodology to extract the time trend in the ``ceteris paribus'' (or fitted) transport cost component, for each the ad-valorem and the additive component.

>>>>

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of export price multiplied by the ad-valorem cost (expressed in Norwegian crowns), i.e. $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results). One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them, allowing us to gain in generality in this respect. Making use of our estimates for 2004 (see Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail} in Appendix \ref{app:more_results}), we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.5\% in air transport. This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases). In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment. \citet{Irrazabal_2015}'s results imply that, for an export price multiplied by the ad-valorem cost of 100 USD, 6 USD are paid in additive trade costs. For the same export price multiplied by the ad-valorem cost of 100 USD, our results yield the value of 2.8 (1.5) USD paid in additive transport costs for maritime (air) transport.\footnote{Also notice that, in both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$. Given our estimates for 2004 ($x = 0.028$ and $0.015$ in vessel and air transport respectively), we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 2.7\% and 1.4\% in each transport mode respectively.} This comparison of results suggests that between 47\% (in Vessel) and 25\% (in Air) and   additive trade costs are attributable to international transport. On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs.\smallskip

Given the time coverage of our database, one may then wonder whether this important part of additive costs we put forward, is a long-lasting, ie structural phenomenon. We investigate this point by reporting the share of additive costs in total transport costs on a yearly basis, in both Air and Vessel transport, based on the weighted-mean values $\widehat{t}/\widetilde{p}$ and $\widehat{\tau}$ (in similar terms as those reported in Tables \ref{tab:result_air_3d_detail} and \ref{tab:result_ves_3d_detail}. The corresponding values reported in detailed tables are available on the Online Appendix), as displayed in Figure \ref{fig:part_cout_additif}. Precisely, each yearly value is obtained as: $\frac{\frac{\widehat{t}}{\widetilde{p}}}{\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}}$, expressed in percent of the export price $\widetilde{p}$ (at the 3-digit level).

\begin{figure}[htbp]
\caption{Share of additive costs (3-digit classification)}
\label{fig:part_cout_additif}
\begin{center}
\includegraphics[width=12cm, height=7cm]{part_cout_additif.pdf}
%{\footnotesize {OECD data} }
\end{center}
\end{figure}

In both transport modes, the additive component appears of sizeable importance throughout the period. Confirming the results obtained in Table \ref{tab:summary_results}, this suggests that additive costs are neither negligible in magnitude nor an erratic phenomenon. By contrast, they represent an sizeable and structural dimension of international transport costs. If the share of additive costs in total costs roughly amounts to 48\% (42\%) in Vessel (Air) transport on average, one can yet notice that it also substantially varies over time, from a minimum of 30\% to a maximum of 60\% depending on the year considered. In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs trend patterns.

@@

In this section, we investigate the trend patterns of international transport costs over time, by exploiting the time dimension of our database. This focus is shared with part of the international trade literature, such as \cite{Lafourcade_Thisse}, \cite{hummels2007} and \cite{Behar_Venables}. Specifically, in the same spirit as \cite{hummels2007} we investigate the underlying sources behind the downward trend of international transport costs, by identifying the respective roles of the reduction in ``pure'' transport costs and changes in trade composition effects.

@@

As a first step, Figure \ref{fig:Trends_in_TC} displays the evolution of total transport costs over the period, summing the additive and the ad-valorem estimated costs, expressed in percentage of the fas price,\footnote{Precisely, we use the yearly values of the estimated transport costs components averaged over the country/product dimensions (at the 3-digit level), corresponding to the numbers reported for each year in detail in the Online Appendix. Each yearly value is obtained as: $\widehat{\tau}-1+\frac{\widehat{t}}{\widetilde{p}}$.} by transport mode for each year between 1974 and 2013.

\begin{figure}[htbp]
\caption{Trend in estimated transport costs (Yearly mean value)}
\label{fig:Trends_in_TC}
\begin{center}
\includegraphics[width=14cm, height=8cm]{Figure1_Trend_of_totalTC_bymode.pdf}
\end{center}
\end{figure}

As reported in Figure \ref{fig:Trends_in_TC}, both air and vessel shipping exhibit a downward trend in overall transport costs since 1974, by -2.1\% per year for mean air transport costs and -2.0\% per year for mean ocean transport costs, implying a 50\% decrease in Air and a 60\% decrease in Vessel over the period.\footnote{One may be puzzled for by the high magnitude of estimates for the beginning of the period (until 1980 approximately) for ocean transport. \citet{hummels2007} finds similar outcomes on tramp prices indexes, and suggests the oil shock as a likely culprit, in a context where technological progress was quicker in aviation than in vessel, allowing a better dampening of oil shocks on air freight rates. In a related manner, one may worry that the strong decrease in transport costs documented in Figure \ref{fig:decomp_TC_3d} springs from high oil-shock related transport costs in 1974. However, computing the time trends from 1980 does not dramatically change the picture. The yearly trend from 1980 is -2\% for mean air transport costs and -1.6\% for mean vessel transport costs. We thus choose to exploit the whole time dimension of our database by taking 1974 as starting date of our time trend analysis.} On US data, \citet{Hummels_1999} obtains that overall transport cost declined from 6 \% to 4 \% of the import value between 1974 and 1996. For the same years, we obtain a total decrease from 6.9 to 4.2\% in terms of the export price, on average for air shipping, and from 9.8\% to 4.8\% for ocean shipping. Overall, our results display trends that are close to those reported by \citet{Hummels_1999}.\smallskip

Before making any definite statement about this though, it is worth emphasizing that the time trend of international transport costs depends on both the evolution of per product and per partner (``pure'') transport costs and the evolution of the composition of trade. Total transport costs may thus have decreased over time because the share of neighboring countries in total US trade or the share of goods cheaper to transport has increased, independently of any change in transport costs \textit{per se}. As emphasized by \citet{Hummels_1999} or \cite{hummels2007}, it is hence necessary to eliminate the composition effects of trade flows to isolate the evolution of pure international transport costs. Specifically, Hummels's (2007) results\nocite{hummels2007} underlie the view that trade composition effects partially compensate the reduction in pure transport costs, thereby attenuating the downward trend in overall (observed) transport costs since 1974. Our own results tend to challenge this view, as we further show.

@@

\paragraph{Challenging the $\beta$'s constancy assumption} As reported in Figure \ref{fig:part_cout_additif} (based on the yearly estimates of transport costs averaged over the country/sector dimension), the share of additive costs is rather varying over time, in both transport modes. In light of this, one may wonder about the constancy of the share of additive costs along the pair product/origin country hypothesized by \cite{hummels2007}. We investigate this point deeper by reporting the histogram of the distribution of the share of additive costs in total costs. To do so, we start from Equation (\ref{eq:base_estimee}) with the trade costs measure on the left-hand side, from which we deduce the formulae to get the elasticity of transport costs to the export price, i.e. $\beta$, according to:
\begin{eqnarray*}
\beta_{ikt} &\equiv& \frac{\partial TC_{ikt}}{\partial \widetilde{p}_{ikt}}\frac{\widetilde{p}_{ikt}}{TC_{ikt}} \\
&=& \frac{\widehat{t}_{ikt}/\widetilde{p}_{ikt}}{\widehat{t}_{ikt}/\widetilde{p}_{ikt}+\widehat{\tau}_{ikt}-1},
\end{eqnarray*}

@@

Estimation driven in Section \ref{sec:results_decomposition} provides us with the additive and ad-valorem measures of international transport costs (Equation (\ref{eq:model_IetA})), that vary over time, product and origin country (i.e., $\widehat{t}_{ikt}$ and $\widehat{\tau}_{ikt}$). Starting from these values, for each additive and multiplicative cost component (by transport mode) we extract the pure transport cost measure by the mean of a time fixed effect. Precisely, we extract the changes over time in the pure transport cost dimension by assuming a composition of trade flows by country partner and product that is constant throughout the period, and equal to the one observed in 1974. We conduct the same analysis on a overall transport costs measure, built by agglomerating the two estimated components (additive and iceberg) in a unified measure of transport costs.\footnote{Note that the unfitted total transport costs are virtually the same as those reported in Figure \ref{fig:Trends_in_TC}, but reported from another perspective (basis 100 in year 1974).} The objective is to obtain six time series, all built as indices with the reference value 100 in 1974: Three time series for the unfitted transport costs measures $\{\Gamma^{add, raw}_t, \Gamma^{adv, raw}_t, \Gamma^{tc, raw}_t \}$  (additive, ad-valorem and total respectively), and three series for the fitted values $\{\Gamma^{add}_t, \Gamma^{adv}_t, \Gamma^{tc}_t \}$ (additive, ad-valorem and total cost respectively).

@@

\paragraph{Obtaining the unfitted total transport cost index} For each transport mode, we build the total transport cost series based on Equation (\ref{eq:base_estimee}) according to:
$$\widehat{tc}^{raw}_t= \widehat{\tau}^{adv}_t -1 + \frac{\widehat{t}_t}{\widetilde{p}_t}$$
\noindent where $\widehat{\tau}^{adv}_t$ and $\widehat{t}_t$ are the average values over the country/sector dimension estimated (conditional on a year and a transport mode), as explained in Section \ref{sec:data_method}. Recall that $\widehat{\tau}^{adv}_t-1$ measuring the ad-valorem transport cost component and $\frac{\widehat{t}_t}{\widetilde{p}_t}$ the additive component, both expressed in percentage of the fas price. We then transform this value in an index with basis year 1974, applying a similar formula as above (by transport mode):
$$\Gamma^{tc, raw}_t = 100\frac{\widehat{tc}^{raw}_t -1 }{\widehat{tc}^{raw}_{1974}-1}$$

@@

\subsection{Characterizing the time trends in transport costs}
Figure \ref{fig:totalTC_compeffects_excl} reports the results for all types of goods.\footnote{In the Online Appendix, we report the results at a more disaggregated level, distinguishing between primary and manufacturing goods.}\textbf{Et alors? Faire un commentaire et mettre l'annexe en ligne} In panels (a), (b) and (c) we report the time changes of the ad-valorem costs, the additive costs and the total costs respectively, for Air transport (starting from the reference value 100 in 1974). Panels (d), (e) and (f) report the results for Vessel. In each panel, we report the evolution of transport costs for both the unfitted (plain blue line) and the fitted (dotted red line) measures.

@@

Three main results emerge from Figure \ref{fig:totalTC_compeffects_excl}. First, in accordance with Figure \ref{fig:decomp_TC_3d}, we find that international transport costs have substantially decreased over the period, for both transport modes (Panels (c) and (f)). International transport costs were reduced by 50\% between 1974 and 2013 in air shipping, and by 60\% in maritime shipping. This stands in line with the related literature (see \citealp{Hummels_1999}, \citealp{Lafourcade_Thisse}). Second, the magnitude of the decrease is roughly of the same order for both for the ad-valorem and the additive components (Panels (a), (b), (d) and (e)). Further, the transport cost reduction is much smoother in maritime transport, while air transport shows more volatility in the trend pattern, in particular in the 1980s and the 2000s. 
Third, and most importantly, we find that composition effects do not play a major role in accounting for the time trend of overall transport costs. Inspecting the panels of Figure \ref{fig:totalTC_compeffects_excl} for Air transport (panels (a) to (c)), we do not find much evidence of a substantial trend difference between the unfitted transport costs measure and the pure transport costs. For all three series (additive, ad-valorem and overall transport costs), the dotted and the plain line follow closely, almost every year throughout the period. Air transport costs were reduced by 50\% between 1974 and 2013, and this is mainly attributable to a reduction in the transport costs ``per se''. Composition effects are more pronounced in vessel transport, for all three series (Figure \ref{fig:totalTC_compeffects_excl}, panels (d) to (f)), where they amplify the reduction in ``pure'' transport costs (the plain line being lower than the dotted line). Considering the raw series (plain line), maritime transport costs have decreased by 60\% over the period, which can be decomposed in a 50\% decrease in transport costs ``per se'' (dotted line), and a 10\% reduction that comes from composition effects (the difference of the two). This is particularly the case for the multiplicative component (panel (d)).\smallskip

@@

These results stand in sharp contrast with \citet{hummels2007}, who obtains that trade composition effects do matter, as they partially offset the reduction in the pure transport costs for both air and maritime shipping. If anything, we find the opposite result here: Trade composition effects do not matter much in accounting for the downward trend of observed international transport costs, and when they do, they tend to amplify (rather than reduce) the downward trend pattern. This drives us to investigate this difference further. As we explain in more details in Appendix \ref{app:comp-effects}, our empirical strategy differs from \cite{hummels2007} in one main dimension. Our characterization of the time trends in transport costs starts from our estimates of both the additive and the ad-valorem components (obtained in Section \ref{sec:results_decomposition}), as well as for the overall transport cost (rather than the actual cif-fas price gap). As direct consequence, our methodology lets the ratio between the additive and the ad-valorem components of transport costs vary over the three time-product-partner country dimensions. %By contrast, \cite{hummels2007} does not take into account the changes in the additive transport cost component, attributing this to a change in the composition of the bundle over time (per country-commodity).
This turns out to be of primary importance in the disentangling in the time trend of transport costs, between what comes from the trade composition effects and what comes from changes in the pure transport cost dimension, as we show below.

To establish this point clearly, we replicate the method adopted by \cite{hummels2007} exposed above on our database (which is the same as his until 2004). The results are reported in Figure \ref{fig:comp_effects_as_in_Hummels}. The dotted line labeled ``expenditure/import value'' represents the unfitted measure of transport costs ($TC_t$ in the above terminology) and the plain line labeled ``fitted ad-valorem rate'' is the measure of ``pure'' transport costs, i.e. composition effects excluded ($\widehat{TC}_t$).

@@

These results stand in sharp contrast with the ones obtained with our methodology and reported in Figure \ref{fig:totalTC_compeffects_excl}. Applying \citeh{hummels2007} method, we find that the composition effects tend to partly offset the decrease in transport costs in both air and vessel shipping, as the downward trend is of higher magnitude for the fitted than the unfitted rate, especially for Vessel.\footnote{Unsurprisingly, this stands in accordance with \citeh{hummels2007} results, see his Figures 5 (for Air) and 6 (for Vessel).} This result is overturned when we allow for more flexibility in the role of the additive component, as pointed in Figure \ref{fig:totalTC_compeffects_excl}, panels (c) and (f).

Assuming a varying share of the additive component over time, product and country partner indeed modifies the decomposition of the trend reduction of transport costs between the one attributable to trade composition effects and the reduction in the ``ceteris paribus'' transport costs. In both air and vessel transports, we thus find that this last dimension is the main driver of the reduction of international transport costs observed over time, in particular in air transport. Complementing the findings of Section \ref{sec:results_decomposition}, these results point out the importance of integrating the additive dimension of international transport costs, here in view of characterizing their time trends.

\subsection{Robustness to the weighting scheme \label{sec:robustness}}

As discussed in Appendix \ref{app:comp-effects}, our methodology also differs from \citet{hummels2007} with respect to the weighting scheme used to obtain the evolution of the pure transport costs over time. In this section, we provide a robustness test to this difference of treatment.\smallskip

When aggregating the trade cost measure over the product/country ($i,k$) dimension, \cite{hummels2007} takes the unweighed average value over the $i,k$ dimension, which implicitly attributes a weight equal to 1 to each flow. Our benchmark specification  (at the root of Figure \ref{fig:totalTC_compeffects_excl}) proceeds differently, as each flow is weighted by its relative value in total trade flows observed in 1974. In this section, we check that this difference of treatment is not responsible for the difference of results relative to the role of the trade composition effects, that we rather attribute to the modeling of the additive component.

@@

Our results could be extended in two main ways. On the empirical side, one may want to go deeper in the structural determinants of (pure) transport costs, i.e. identify the respective roles of handling costs, insurance and freight at the root of the gap between export and import prices. On the theoretical side, our results can be used to explore the role of additive costs in shaping international trade flows (in an international trade theory perspective) and in affecting the international transmission of business cycles. This is left for further research.

@@

@@

In this section, we detail our methodology to extract the time trend in the pure (or fitted) transport cost component, for each the ad-valorem and the additive component.
